\sysfull

### 2018-08-04 paper/asplos19/abstract.tex
partial commit

%Old version of abstarct

%Massively deployed Internet of Things (IoT) must be energy neutral. This can only be achieved when the IoT is battery-free and the energy to operate the IoT comes from ambient energy sources. However, relying on the highly varying, unpredictable and intermittent harvested energy is at odds with the core assumption that program executions require stable and uninterrupted energy supply. The core programming model for the intermittently-powered computation is a discrete-like one that alternates between progressing and protecting the execution progress against power interrupts. Generally, intermittent computation models can be classified into (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based programming, where a programmer splits the code into small idempotent sections. Despite the superiority of task-based systems over checkpointing, task-based programming is completely oblivious to the energy conditions. Simply, it cannot increase or decrease the size of the task on the fly (the moment of commit to the non-volatile memory) depending on how long the energy burst is. To address this challenge, this paper introduces \sys, an \emph{adaptive} task-based execution model that does not require any dedicated hardware to support task adaptability, which at the same time is \emph{efficient}. \sys programs tune the frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period off the programmer's shoulders. \sys defaults to progress on a multi-tasks scale, when energy permits (\emph{efficiency}), and on a sub-task scale when needed (\emph{guarantee}). Furthermore, \sys's \emph{memory virtualization} gains \sys full data protection against runtime triggered and compile-time defined memory accesses. Results show that \sys is more efficient than the current state-of-art execution model by up to a factor of two and able to progress where its counterparts fail.
%
%Alexei's version of abstract
%
% Embedded devices in the Internet of Things gain new deployment opportunities when their hardware is free of batteries and powered by ambient energy. Energy harvested from the environment, however, is unpredictable and can only power a device intermittently. In the intermittent execution model, computation is frequently interrupted by power failures, and to make forward progress, the system must save intermediate program state into non-volatile memory. Task-based programming models that execute at the granularity of statically-defined tasks, offer an efficient alternative to checkpointing. Existing task-based systems, however, fix their task size at compile time and operate without any regard to the changing energy conditions at runtime. When tasks are sized without accounting for available energy, state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate. To address these challenges, we propose \sys, an \emph{adaptive} task-based execution model that progresses on a multi-tasks scale, when energy permits (\emph{efficiency}), and preserves the computation progress on a sub-task scale when needed. \sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism. Our evaluation on a real energy-harvesting device demonstrates the benefits of adaptive task coalescing and downscaling, and highlights \sys's performance improvements over a state-of-art task-based system.
%
Freeing embedded devices from batteries is a key factor in developing the energy neutral Internet of Things. 
Energy harvesting, however, is unpredictable and can only power a device intermittently. Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve the execution progress of a program. Despite the superiority of the task-based approach, however, it fixes the task size at compile time and operates obliviously to changing energy conditions. When tasks are sized without accounting for available energy, the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.% When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate. 
To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model. \sys progresses on a multi-tasks scale when energy permits and preserves the computation progress on a sub-task scale if necessary. \sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism. Our evaluation shows that \sys is $2X$ faster than the state-of-the-art system in most cases and is able to progress where static systems fail.

>>>>

Freeing embedded devices from batteries is a key factor in developing the energy neutral Internet of Things. 
Energy harvesting, however, is unpredictable and can only power a device intermittently. Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve the execution progress of a program. Despite the superiority of the task-based approach, however, it fixes the task size at compile time and operates obliviously to changing energy conditions. When tasks are sized without accounting for available energy, the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model. \sys progresses on a multi-tasks scale when energy permits and preserves the computation progress on a sub-task scale if necessary. \sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism. Our evaluation shows that \sys is $2X$ faster than the state-of-the-art system in most cases and is able to progress where static systems fail.

### 2018-08-04 paper/asplos19/background.tex
partial commit

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{Subcaption}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{Subcaption}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{Main caption.}
    \label{fig:war}
\end{figure}

@@

A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.

Checkpointing volatile state alone do not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Especially, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em
reads}~\cite{ratchet,dino,alpaca}\TODO{figure 2 dependent text ---see
Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
inconsistent in an intermittent execution using the cyclic redundancy check
(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
maintain data consistent in the presence of WAR dependency is necessary for
correct execution.}

>>>>

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{Incorrect execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependence example. \texttt{NV} marks non-volatile variables,
    while \texttt{checkp()} is a checkpoint of volatile state.}
    \label{fig:war}
\end{figure}

@@

A power failure clears volatile state (e.g., registers and SRAM) while
non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
to a prior point in the execution: by default, to the beginning of {\tt
main()}. Early intermittent systems preserved progress by periodically
checkpointing volatile execution context to non-volatile
memory~\cite{mementos}, sometimes using hardware
support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. 
%
Checkpointing volatile state alone does not ensure data consistency when the
system can directly manipulate non-volatile memory~\cite{mspcdino}.
Precisely, data can get inconsistent when code includes a
\emph{write-after-read} (WAR) dependence between operations that manipulate
non-volatile memory, because it allows {\em writes} from a failed attempt to
persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
Figure~\ref{fig:war} illustrates how state can become
inconsistent in an intermittent execution using a simple example of an average
operation over an array of integers.
The variable \texttt{sum} introduces the WAR, being read and written sequentially
by the increment operation, thus being erroneously incremented when a power
failure occurs before updating the index \texttt{i}.

### 2018-08-04 paper/asplos19/evaluation.tex
partial commit

% \begin{wrapfigure}{t!}{0.5\textwidth}
\begin{figure}
	%\begin{figure}
	\centering
	\includegraphics[width=0.5\columnwidth]{figures/coalEfficiency}
    \caption{Breakdown of overhead for the proposed coalescing strategies
relative to the baseline without coalescing.
NC: no coalescing, EO: energy-oblivious coalescing, HG: history-guided coalescing, WHG:
energy- and task-aware coalescing. All coalescing strategies reduce total
overhead and maximize useful work. HG and WHG perform better than EO.}
	\label{fig:overallOverheadBreakdown}
\end{figure}
% \end{wrapfigure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.49\columnwidth]{figures/overallOverhead}
	\includegraphics[width=0.49\columnwidth]{figures/memAccess}
	\caption{\sys kernel overhead: a breakdown of time spent in
the application and in the runtime (left) and a breakdown of protected memory
accesses into three types with different costs. NC: no coalescing, HG:
history-guided coalescing, WHG: weighted history-guided coalescing.}
	\label{fig:coalEfficiency}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}%
	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
    \caption{\sys's coalescing performance. Left: Application execution time
with coalescing strategies (HG and WHG) normalized to the execution time
without coalescing. Right: Execution time with the WHG variant of \sys,
normalized to the execution time with Alpaca, as a function of distance to the
energy source.}
	\label{fig:coalescing}
	\label{fig:runtime}
\end{figure}

%\begin{wrapfigure}{t!}{0.5\textwidth}
%	%\begin{figure}
%	\centering
%	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}
%	\caption{Coalescing strategies (HG and WHG) run time per application on intermittent power (RF generator at 15\,cm from the WISP), compared to \sys without coalescing (NC). Coalescing provides significant improvements: from 0.25 (\textbf{ar}) up to 0.70 (\textbf{sort}) compared to a non-coalescing system.}
%	\label{fig:coalescing}
%	%\end{figure} 
%\end{wrapfigure}
%
%\begin{wrapfigure}{t!}{0.5\textwidth}
%	%\begin{figure}
%	\centering
%	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
%    \caption{Application execution time with the WHG variant of \sys,
%normalized to the execution time with Alpaca, as a function of distance to the
%signal generator antenna. \sys outperforms Alpaca on the majority but not all
%of the benchmarks.}
%	\label{fig:runtime}
%%\end{figure}
%\end{wrapfigure}

%\textbf{Benefits of Adaptive Tasks.}

>>>>

% \begin{wrapfigure}{t!}{0.5\textwidth}
\begin{figure}
	%\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
    \caption{Breakdown of overhead for the proposed coalescing strategies
relative to the baseline without coalescing.
NC: no coalescing, EO: energy-oblivious coalescing, HG: history-guided coalescing, WHG:
energy- and task-aware coalescing. All coalescing strategies reduce total
overhead and maximize useful work. HG and WHG perform better than EO.}
	\label{fig:overallOverheadBreakdown}
\end{figure}
% \end{wrapfigure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/overallOverhead}
	\includegraphics[width=\columnwidth]{figures/memAccess}
	\caption{\sys kernel overhead: a breakdown of time spent in
the application and in the runtime (left) and a breakdown of protected memory
accesses into three types with different costs. NC: no coalescing, HG:
history-guided coalescing, WHG: weighted history-guided coalescing.}
	\label{fig:coalEfficiency}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/coalStrategies}%
    \caption{\sys's coalescing performance. Left: Application execution time
with coalescing strategies (HG and WHG) normalized to the execution time
without coalescing.}
	\label{fig:coalescing}
\end{figure}

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
    \caption{\sys's execution time, normalized to the execution time 
    with Alpaca, as a function of distance to the energy source 
    measured in meter.}
	\label{fig:runtime}
\end{figure}
